---
title: "p_8105_hw2_ep2899"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df=
  read_xlsx("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx")
```

Specify
```{r}
trashwheel_df=
read_xlsx(
  "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
  sheet= "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

```{r}
median (pull(trashwheel_df, sports_balls), na.rm=FALSE)
```

Read the precipitation data for 2018 and 2017!

```{r}
 precip_2018=
    read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx" ,
     sheet = "2018 Precipitation",
     skip = 1
    ) %>% 
   janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year=2018) %>% 
  relocate(year) 
```

```{r}
colSums(precip_2018, na.rm=TRUE)
```
     
 
```{r}
 precip_2017=
  read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx" ,
     sheet = "2017 Precipitation",
     skip = 1
  ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

precip_2017=

Now combine annual precipitation.
 
```{r}
precip_df=
  bind_rows(precip_2017, precip_2018)
```
 
```{r}
month_df=
  tibble(
    month = 1:12,
    month_name = month.name
    )

precip_df=
  bind_rows(precip_2018, precip_2017)

left_join (precip_df, month_df, by= "month")
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of 'r nrows(trashwheel_df)' rows in our final dataset. Additional data sheets include month precipitation data.
Total precipitation in 2018 is 70.33, based on code in line 55
Median number of sports balls in a dumpster in 2017 is 8, based on code in line 36 above

## Problem 2
## Load tidyverse
```{r setup}
library(tidyverse)
```

## Read CSV
```{r}
nyctransit_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
nyctransit_data = janitor::clean_names(nyctransit_data)
```

##clean data
```{r}
data_clean= select(nyctransit_data, -division, -staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location) 
data_clean
```


##convert character to logical
```{r}
data_clean= data_clean %>%  
mutate(entry=ifelse(entry=="YES", TRUE, FALSE))
```

## short paragraph
This is a csv file data of NYC transit info containing 1868 rows x 20 columns, with information regarding line, station name, station latitude, station longitude, routes served, entry, vending, entrance type, and ADA compliance 
This data is not yet tidy

##distinct stations
```{r}
data_clean %>% 
distinct(line, station_name)%>% 
count()
```
There are 465 distinct stations, by line and station_name

##filter ADA
```{r}
filter(data_clean, ada==TRUE) 
```
There are 458 stations which are ADA compliant

##proportion entrance
```{r}
data_clean= data_clean %>%  
mutate(vending=ifelse(vending=="YES", TRUE, FALSE))
```

```{r}
select (data_clean, entrance_type, exit_only, entry, vending) %>%  
filter(vending==FALSE) %>% 
  filter(entry==TRUE)
```
69/1868 or 3.7% of station entrances and exits without vending allow entrance

##reformat



##Problem 3
pols month
```{r}
polsmonth_df = read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year", "month", "day")) %>% 
  mutate(president = prez_gop + prez_dem,
         year = as.numeric(year),
         day = as.numeric(day),
         month = as.numeric(month)
         )
```

```{r}
polsmonth_df=
select(polsmonth_df, year, month, gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, rep_dem)
```

```{r}
polsmonth_df_mut=
  tibble(
    month = 1:12,
    month_name = month.name
    ) 
```

```{r}
polsmonth_df_mname=
left_join (polsmonth_df, polsmonth_df_mut, by= "month")
```

snp
```{r}
snp_df = read_csv("./data/snp.csv") %>% 
janitor::clean_names() %>% 
  separate(date, c("month", "day", "year")) %>% 
   mutate (month = as.numeric (month), day = as.numeric(day), year = as.numeric(year))
```

```{r}
snp_df_mname=
  left_join (snp_df, polsmonth_df_mut, by= "month")
```

```{r}
snp_df_mname=
  select(snp_df_mname, -month)
```

unemployment
```{r}
unemp_data = read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "rate"
  )
```


Merge snp into pols
```{r}
snp_pols=  
left_join (snp_df_mname, polsmonth_df_mname, by= "month_name") %>% 
janitor::clean_names()
```

Merge unemp
