---
title: "p_8105_hw1_ep2899"
output: github_document
---

## Problem 2
## Load tidyverse
```{r setup}
library(tidyverse)
```

## Read CSV
```{r}
nyctransit_data = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
nyctransit_data = janitor::clean_names(nyctransit_data)
```

##clean data
```{r}
data_clean= select(nyctransit_data, -division, -staffing, -staff_hours, -ada_notes, -free_crossover, -north_south_street, -east_west_street, -corner, -entrance_latitude, -entrance_longitude, -station_location, -entrance_location) 
data_clean
```


##convert character to logical
```{r}
data_clean= data_clean %>%  
mutate(entry=ifelse(entry=="YES", TRUE, FALSE))
```

## short paragraph
This is a csv file data of NYC transit info containing 1868 rows x 20 columns, with information regarding line, station name, station latitude, station longitude, routes served, entry, vending, entrance type, and ADA compliance 
This data is not yet tidy

##distinct
```{r}
data_clean %>% 
distinct(line, station_name)%>% 
count()
```
There are 465 distinct stations, by line and station_name

##filter
```{r}
filter(data_clean, ada==TRUE) 
```
There are 458 stations which are ADA compliant

```{r}
data_clean= data_clean %>%  
mutate(vending=ifelse(vending=="YES", TRUE, FALSE))
```

```{r}
select (data_clean, entrance_type, exit_only, entry, vending) %>%  
filter(vending==FALSE) %>% 
  filter(entry==TRUE)
```
69/1868 or 3.7% of station entrances and exits without vending allow entrance  

